---
title: "Datpiff"
author: "Rob Davis"
date: "11/2/2020"
output: 
    html_document: default
    html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(VIM)
library(reshape2)
library(scales)

options(dplyr.summarise.inform = FALSE)
```

## Import Data
```{r}
raw.data_0 <- read.csv('../datpiff_2020-11-03_imputed.csv')
```

#### Let's take a quick look at the data
```{r}
summary(raw.data_0)
```
#### Rename the first column which is screwed up for some reason
#### Drop the description field for now, which doesn't have much significance or usability
#### convert release_date to a date
#### Change these (no/yes) fields to (0/1) (buy/download/streaming)_enabled
#### Drop mixtapes with 0 tracks
#### Drop mixtapes where listens > views or downloads > views (these are few and likely erroneous or fraudulent)
#### All mixtapes have more than 0 listens, therefore 0 views is an error and should be imputed (set as NA for now)
```{r}
raw.data_1 <- raw.data_0 %>% 
  rename(., added_by = 'ï..added_by') %>% 
  select(., !c(description)) %>% 
  mutate(., release_date = mdy(release_date),
         buy_enabled = ifelse(buy_enabled=='yes',1,0),
         download_enabled = ifelse(download_enabled=='yes',1,0),
         streaming_enabled = ifelse(streaming_enabled=='yes',1,0),
         views = ifelse(views==0,NA,views)) %>% 
  filter(., tracks > 0, views > listens)

summary(raw.data_1)
```
#### Drop Streaming Enabled because it's always 1
```{r}
raw.data_1 <- select(raw.data_1, !streaming_enabled)
```

#### Investigating NA downloads
```{r}
raw.data_1 %>%
  filter(., download_enabled==1, is.na(downloads))
```
#### Upon investigation of the site, these mixtapes have errors on their pages where downloads failed to load, along with views.

## Imputation Efforts
#### Mixtapes with downloads enabled and NA downloads
```{r}
dl_able <- filter(raw.data_1, download_enabled==1)
no_dl_able <- filter(raw.data_1, download_enabled==0)

dl_able_impute <- VIM::kNN(dl_able, variable = 'downloads', k = 5, imp_var = FALSE)

raw.data_2 <- full_join(no_dl_able, dl_able_impute)

summary(raw.data_2)
```

#### Impute Mixtapes with 0 views
```{r}
data_no_dl <- raw.data_2 %>% 
  mutate(., views = ifelse(views==0,NA,views)) %>% 
  select(., !(downloads))

data_no_views <- raw.data_2 %>% 
  select(., !(views))

impute_views_no_dl <- VIM::kNN(data_no_dl, variable = 'views', imp_var = FALSE, k = 5)

raw.data_3 <- full_join(impute_views_no_dl, data_no_views)

data <- raw.data_3
```

## Adding Features: 
#### listens_per_view, downloads_per_view, release_month (in YYYY-MM format), release_week
```{r}
data <- data %>% 
  mutate(., 
         listens_per_view = listens / views,
         downloads_per_view = downloads / views,
         release_year = year(release_date),
         release_month = floor_date(release_date, unit = "month"), 
         release_week = floor_date(release_date, unit = 'week'))
```

## Some analysis:

### Correlations
```{r}
cor_dl <- data %>% 
  filter(., download_enabled == 1) %>% 
  select(., !c(added_by, artist, banner, host, release_date,release_month, release_year, release_week, title,
               download_enabled, buy_enabled)) %>%
  cor(.)

cor_no_dl <- data %>% 
  filter(., download_enabled == 0) %>% 
  select(., !c(added_by, artist, banner, host, release_date,release_month, release_year, release_week, title, 
               download_enabled, downloads, downloads_per_view)) %>%
  cor(.)

cormap_to_map <- function(cormap)
{
  cormap[upper.tri(cormap)] <- NA
  cormap <- melt(cormap, na.rm = T)
  return(cormap)
} 

cormap_dl_plot <- cormap_to_map(cor_dl)

cormap_no_dl_plot <- cormap_to_map(cor_no_dl)

make_heatmap <- function(cormap, title){
  out <- ggplot(cormap, aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile(color = "white") +
    scale_fill_gradient2(low = "gray", high = "orange", 
                         midpoint = 0.5, limit = c(0,1), space = "Lab", 
                         name="Correlation") +
    theme_minimal()+ 
    theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                     size = 12, hjust = 1)) +
    coord_fixed() + 
    ggtitle(title)
  
  return(out)
}

heatmap_dl <- make_heatmap(cormap_dl_plot, "Correlation Heatmap for Mixtapes\nwith Downloads Enabled")

heatmap_no_dl <- make_heatmap(cormap_no_dl_plot, "Correlation Heatmap for Mixtapes\nwith Downloads Disabled")
```

```{r}
heatmap_dl
```
```{r}
heatmap_no_dl
```

#### Grouping Data monthly + Plotting
```{r}
monthly <- data %>% 
  group_by(., release_month) %>% 
  summarise(., count = n(), 
            avg_views = mean(views), med_views = median(views),
            avg_listens = mean(listens), med_listens = median(listens), 
            avg_downloads = mean(downloads, na.rm = T), med_downloads = median(downloads, na.rm = T),
            avg_lpv = mean(listens_per_view), med_lpv = median(listens_per_view),
            avg_dpv = mean(downloads_per_view, na.rm = T), med_dpv = median(downloads_per_view, na.rm = T),
            total_views = sum(views), total_listens = sum(listens), total_downloads = sum(downloads, na.rm = T))
```

```{r}
monthly_stats <- monthly %>% 
  pivot_longer(.,
               count:total_downloads,
               names_to = "stat",
               values_to = "value")

monthly_stats
```

#### Plotting Setup
```{r}
view_stats = c("avg_views", "med_views")
download_stats = c("avg_downloads", "med_downloads")
listen_stats = c("avg_listens", "med_listens")
freq_stats = c("count")
lpv_stats = c("avg_lpv", "med_lpv")
dpv_stats = c("avg_dpv", "med_dpv")
listens_downloads = c("avg_views", "avg_listens", "avg_downloads")
med_l_dl = c("med_views", "med_listens", "med_downloads")
lpv_dpv = c("avg_lpv", "avg_dpv")
totals = c("total_views", "total_listens", "total_downloads")

stat_plot <- function(x){
  monthly_stats %>% 
    filter(., stat %in% x) %>%
    ggplot(., aes(x = release_month, y = value)) +
    geom_line(aes(color = stat)) 
  }

```

### Mixtape upload frequency over time
```{r}
stat_plot(freq_stats) +
  ggtitle("Mixtape Uploads per Month (2010-2020)") +
  ylab("Mixtapes Uploaded") + 
  xlab("Month") +
  theme_minimal() + 
  theme(legend.position = "none") +
  scale_x_date(date_labels = "%b-%y", date_breaks = "year", limits = c(ymd("2010-01-01"),NA)) + 
  scale_y_continuous(breaks = seq(0,140,20), limits = c(0,140))

```

### Upload Frequency of sponsored / exclusive mixtapes
```{r}
paid_monthly <- data %>% 
  mutate(., sponsored = ifelse(banner == "none", "unsponsored", "sponsored")) %>% 
  group_by(., release_month, sponsored) %>% 
  summarise(., count = n(), 
            avg_views = mean(views), med_views = median(views),
            avg_listens = mean(listens), med_listens = median(listens), 
            avg_downloads = mean(downloads, na.rm = T), med_downloads = median(downloads, na.rm = T),
            avg_lpv = mean(listens_per_view), med_lpv = median(listens_per_view),
            avg_dpv = mean(downloads_per_view, na.rm = T), med_dpv = median(downloads_per_view, na.rm = T),
            total_views = sum(views), total_listens = sum(listens), total_downloads = sum(downloads, na.rm = T)) %>% 
  pivot_longer(.,
               count:total_downloads,
               names_to = "stat",
               values_to = "value") 
  #%>%
  #mutate(., banner = factor(banner, levels = c("none", "exclusive", "sponsored")))
  #%>% 
  # select(., !banner)

paid_monthly

paid_uploads_plot <- function(stats, title = NULL){
  paid_monthly %>% 
    filter(stat %in% stats, between(release_month, ymd("2009-06-30"), ymd("2020-12-31"))) %>% 
    ggplot(., aes(x = release_month, y = value, fill = sponsored, color = sponsored)) +
    geom_line() + 
    ggtitle(title) +
    theme_minimal() +
    theme(legend.position = c(.85,.85)) +
    xlab("Month") +
    ylab("Mixtapes Uploaded") + 
    scale_x_date(date_labels = "%b-%y", date_breaks = "year", limits = c(ymd("2010-01-01"),NA)) + 
    #scale_x_date(date_breaks = "2 years", date_labels = "%b-%y") +
    scale_y_continuous(breaks = seq(0,140,20), limits = c(0,140)) +
    scale_color_discrete(name = "Sponsorship Status", labels = c("Sponsored", "Unsponsored"))
  }

paid_uploads_plot(c("count"), title = "Mixtape Sponsorship per Month (2010-2020)")

```

## Total Engagement by Release Date
```{r}
stat_plot(totals) + 
  ggtitle("User Engagement by Release Month (2010-2020)") + 
  xlab("Release Month") +
  ylab("Monthly Total") + 
  theme_minimal() + 
  theme(legend.position = c(.85,.85)) +
  scale_x_date(date_labels = "%b-%y", date_breaks = "2 years", limits = c(ymd("2009-06-30"), NA)) +
  scale_y_continuous(labels = comma, breaks = seq(0, 6E7, 1E7)) +
  scale_color_discrete(name = NULL, 
                       breaks = c("total_views", "total_listens", "total_downloads"), 
                       labels = c("Views", "Listens", "Downloads"))
```

#### Quick peek at that peak month
```{r}
data %>% 
  filter(release_month == ymd("2015-11-01")) %>% 
  arrange(., desc(views)) %>% 
  select(., artist, title, views, listens, downloads)
```

## Average and median mixtape views over time
```{r}
stat_plot(view_stats) + 
  ggtitle('Mixtape Views by Release Month') + 
  xlab("Month") +
  ylab("Views") + 
  theme_minimal() + 
  theme(legend.position = c(.85,.85)) +
  scale_x_date(date_labels = "%b-%y", date_breaks = "2 years", limits = c(ymd("2009-06-30"),NA)) +
  scale_y_continuous(limits = c(0, 1250000), breaks = seq(0,1250000, 250000), label = comma) + 
  scale_color_discrete(name = NULL, labels = c("Monthly Average", "Monthly Median"))

```

## Average and median mixtape listens over time
```{r}
stat_plot(listen_stats) + 
  ggtitle('Mixtape Listens by Release Month') + 
  xlab("Release Month") +
  ylab("Listens") + 
  theme_minimal() + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "2 years", limits = c(ymd("2009-06-30"), NA)) +
  scale_y_continuous(limits = c(0, 550000), labels = comma, breaks = seq(0,55E5,1E5)) + 
  scale_color_discrete(name = NULL, labels = c("Monthly Average", "Monthly Median"))
```

## Average and median mixtape downloads over time
```{r}
stat_plot(download_stats) + 
  ggtitle("Mixtape Downloads by Release Month") + 
  xlab("Release Month") +
  ylab("Downloads") + 
  theme_minimal() + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "2 years", limits = c(ymd("2009-06-30"), NA)) +
  scale_y_continuous(limits = c(0, 200000), labels = comma) + 
  scale_color_discrete(name = NULL, labels = c("Monthly Average", "Monthly Median"))
```

## Average and median mixtape listens per view over time
```{r}
stat_plot(lpv_stats) + 
  ggtitle("Listen per View Rate by Release Date") + 
  xlab("Release Month") +
  ylab("Listens per View") + 
  theme_minimal() + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "2 years", limits = c(ymd("2009-06-30"), NA)) +
  scale_y_continuous(limits = c(0, 0.55)) + 
  scale_color_discrete(name = NULL, labels = c("Monthly Average", "Monthly Median"))
```

## Average and median mixtape downloads per view over time
```{r}
stat_plot(dpv_stats) + 
  ggtitle("Download per View Rate by Release Month") + 
  xlab("Release Month") +
  ylab("Monthly Average") + 
  theme_minimal() + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "2 years", limits = c(ymd("2009-06-30"), NA)) +
  scale_y_continuous(limits = c(0, 0.25)) + 
  scale_color_discrete(name = NULL, labels = c("Monthly Average", "Monthly Median"))
```

### Monthly Stats
```{r}
stat_plot(listens_downloads) +
  scale_x_date(limits = c(ymd("2013-06-30", NA)), date_breaks = "2 years", date_labels = "%b-%Y") + 
  scale_y_continuous(limits = c(0, 1250000), labels = comma, breaks = seq(0,1200000,200000)) + 
  ggtitle("Average Mixtape by Release Month") + 
  xlab("Release Month") +
  ylab("Monthly Average") + 
  theme_minimal()+
  theme(legend.position = c(.85,.85)) +
  scale_color_discrete(name = NULL, breaks = c("avg_views", "avg_listens", "avg_downloads"),
                       labels = c("Views", "Listens", "Downloads"))

```

# Median
```{r}
stat_plot(med_l_dl) +
  scale_x_date(limits = c(ymd("2009-06-30", NA)), date_breaks = "2 years", date_labels = "%b-%Y") + 
  scale_y_continuous(limits = c(0, 300000), labels = comma, breaks = seq(0, 3E5, 5E4)) + 
  ggtitle("Median Mixtape by Release Month") + 
  xlab("Release Month") +
  ylab("Monthly Median") + 
  theme_minimal() +
  scale_color_discrete(name = NULL, breaks = c("med_views", "med_listens", "med_downloads"), 
                       labels = c("Views", "Listens", "Downloads"))

```

## 'Per view' Rates together
```{r}
per_view <- stat_plot(lpv_dpv) + 
  #geom_smooth(aes(color = stat), se=T, alpha = 0.2, size = 0.4) +
  ggtitle("Rates of Engagement per View") + 
  xlab("Release Month") +
  ylab("Rate") + 
  theme_minimal() + 
  theme(legend.position = c(.8,.9)) +
  scale_x_date(date_labels = "%b-%y", date_breaks = "2 years", limits = c(ymd("2009-06-30"), NA)) +
  scale_y_continuous(limits = c(0, 0.6), breaks = seq(0, 0.60, 0.05)) +
  scale_color_discrete(name = NULL, breaks = c("avg_lpv", "avg_dpv"), labels = c("Listen Rate", "Download Rate"))

per_view
```

## Median Listens by Download Enabled status
```{r}
data %>%
  group_by(., release_month, download_enabled) %>% 
  summarise(med_listens = median(listens)) %>% 
  ggplot(., aes(x = release_month, y = med_listens, color = as.factor(download_enabled))) + 
  geom_line() + 
  ggtitle("Median Monthly Listens by Download Enabled Status") +
  xlim(ymd("2017-01-01",NA)) + ylim(0,180000) + 
  ylab("Median Monthly Listens") + 
  xlab("Release Month") +
  scale_color_discrete(name = NULL, labels = c("Downloads Disabled", "Downloads Enabled"))
```

### Summarising Data by Artist percentile
```{r}
by_artist <- data %>% 
  group_by(., release_year, artist) %>% 
  summarise(., mixtapes = n(), download_enabled = sum(download_enabled), buy_enabled = sum(buy_enabled), 
            total_views = sum(views), avg_views = mean(views), 
            total_listens = sum(listens), avg_listens = mean(listens),
            total_downloads = sum(downloads, na.rm = T), avg_downloads = mean(downloads)) %>% 
  group_by(., release_year) %>% 
  mutate(., ptile = ntile(desc(avg_views), 20))

annl_artist_ptile <- by_artist %>% 
  select(., release_year, artist, ptile)
  
by_artist

data_ptile <- inner_join(data, annl_artist_ptile)

data_ptile
```

### Download / Buy Rates by ptile
```{r}
data_ptile %>% 
  group_by(., release_month, ptile, download_enabled)
```

### Total listens by annual artist percentiles (in 5%s) for the past 5 years (2015-2020)
```{r}
by_artist %>% 
  filter(., between(release_year, 2015, 2020)) %>% 
  ggplot(., aes(x = ptile, y = avg_listens)) + 
  geom_histogram(stat = "identity")
```

### Histogram
```{r}
hist_data <- data %>% 
  filter(., between(release_year, 2015, 2020)) %>% 
  mutate(., views_bucket = ifelse(views > 1E6, 1E6+1, views))

hist_avg <- mean(hist_data$views)

hist_breaks = seq(0, 10E5, 1E5)
hist_labels = append(seq(0, 900, 100), "1000\nand Greater")

histogram <- hist_data %>% 
  ggplot(., aes(x = views_bucket)) + 
  geom_histogram(breaks = seq(0, 1.05E6, 5E4)) + 
  ggtitle("Frequency of View Counts Among Mixtapes (2015-2020)") + 
  theme_minimal() + 
  ylab("Count") + 
  xlab("Number of Views (Thousands)") + 
  scale_x_continuous(breaks = hist_breaks, labels = hist_labels) + 
  stat_bin(breaks = seq(0, 1.05E6, 5E4), geom="text", aes(label=..count..), vjust=-0.5) 

histogram
```

### Same but log
```{r}
data %>% 
  filter(., between(release_year, 2015, 2020)) %>% 
  ggplot(., aes(x = log(views))) + 
  geom_density() + 
  ggtitle("Density of Views Across Mixtapes (2015-2020)") + 
  theme_minimal() + 
  ylab("Density") + 
  xlab("Number of Views") + 
  scale_x_continuous(label = comma)

```

### Rate of 'Download Enabled' and 'Buy Enabled'
```{r}
options_enabled <- data %>% 
  group_by(., release_month) %>% 
  summarise(., count = n(), download_enabled = sum(download_enabled), buy_enabled = sum(buy_enabled)) %>% 
  mutate(., download_enabled_rate = download_enabled / count, buy_enabled_rate = buy_enabled / count) %>%
  pivot_longer(., count:buy_enabled_rate)

# options_enabled_ptile <- data_ptile %>% 
#   group_by(., release_year, ptile) %>% 
#   summarise(., count = n(), download_enabled = sum(download_enabled), buy_enabled = sum(buy_enabled)) %>% 
#   mutate(., download_enabled_rate = download_enabled / count, buy_enabled_rate = buy_enabled / count) %>%
#   pivot_longer(., count:buy_enabled_rate)
```

```{r}
# options_enabled_ptile%>% 
#   filter(name == "download_enabled_rate", ptile %in% c(1,5,10,15,20)) %>% 
#   ggplot(., aes(x = release_year, y = value)) +
#   geom_line(aes(color = as.factor(ptile))) + 
#   ggtitle("Mixtapes with Purchase and Download Options (2015 - 2020)") +
#   theme_minimal() + 
#   xlim(2015,2020) +
#   ylab("Mixtapes")

```

### Options Enabled Plot
```{r}
options_enabled %>% 
  filter(name %in% c("download_enabled", "buy_enabled")) %>% 
  ggplot(., aes(x = release_month, y = value)) +
  geom_line(aes(color = name, position = "stacked")) + 
  ggtitle("Mixtapes with Purchase and Download Options (2015 - 2020)") +
  theme_minimal() + 
  ylab("Mixtapes") + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "year", limits = c(ymd("2014-06-30"), NA)) +
  scale_color_discrete(name = NULL, labels = c("Purchase Enabled", "Download Enabled")) + 
  xlab("Month")
```

### Options Enabled Rate Plot
```{r}
options_enabled_rates <- options_enabled %>% 
  filter(name %in% c("download_enabled_rate", "buy_enabled_rate"))
  
options_enabled_plot <- options_enabled_rates %>% 
  ggplot(., aes(x = release_month, y = value)) +
  geom_line(aes(color = name)) + 
  ggtitle("Rate of Purchase and Download Options (2015-2020)") +
  theme_minimal() +
  theme(legend.position = c(.85,.85)) +
  ylab("Rate of Appearance") + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "year", limits = c(ymd("2015-01-01"), ymd("2020-12-31"))) +
  scale_color_discrete(name = NULL, labels = c("Purchase Enabled", "Download Enabled")) + 
  xlab("Release Month")

options_enabled_plot
```

```{r}
options_engangement <- options_enabled_rates %>% 
  ggplot(., aes(x = release_month, y = value)) +
  geom_line(aes(color = name)) + 
  ggtitle("Rate of Purchase and Download Options (2015-2020)") +
  theme_minimal() + 
  ylab("Rate of Appearance") + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "year", limits = c(ymd("2015-01-01"), NA)) +
  scale_color_discrete(name = NULL, labels = c("Purchase Enabled", "Download Enabled")) + 
  xlab("Release Month")
```